---
title: "Lab 4"
author: "Christopher Hainzl"
date: "2023-04-18"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
```



```{r}
healthdata <- read_csv("lab4data(1).csv")
healthdata
```
- Weight (lbs) = The measured weight, in pounds, of each subject.
- Height (in.) = The measured height, in inches, of each subject.
- Protein = The protein intake of each subject.
- Carbohydrates = The carbohydrate intake of each subject.
- Calcium = The calcium intake of each subject.
- Gender = The gender of each subject (1 = male, 0 = female)
- Calories = The amount of calories consumed by each subject.


The most appropriate way to begin this analysis would be to set up a linear regression model that predicts a subject's caloric intake based on weight, height, protein, carbohydrates, calcium, and gender. While setting up this model, the assumptions are made
that the model's residuals will be normally distributed
and that the variances are equal. We will also be working with a 95% confidence 
level.

```{r}
# Model 1 - All predictors included
healthmodel1 <- lm(Calories~Weight+Height+Protein+Carbohydrates+Calcium+
gender, data = healthdata)
summary(healthmodel1)
```
- For each unit increase in the subject's weight, their caloric intake increases 
by approximately 1.25 on average. (B1)
- For each unit increase in the subject's height, their caloric intake increases
by approximately 3.01 on average. (B2)
- For each unit increase in the amount of protein consumed by the subject,
their caloric intake increases by approximately 13.43 on average. (B3)
- For each unit increase in the amount of carbohydrates consumed by the subject,
their caloric intake increases by approximately 4.59 on average. (B4)
- For each unit increase in the amount of calcium consumed by the subject,
their caloric intake decreases by approximately 0.24 on average. (B5)
- For the subjects that are males, their caloric intake decreases by
approximately 70.03 on average when compared to females. (B6)

However, based on the p-values associated with each of these predictors, the only
ones that appear to be significant when working with a 95% confidence level are
protein, carbohydrates, and calcium. The p-values associated with each of those
predictors are all less than .05.

Therefore, we should try removing each of the non-significant predictors one at
a time and determine if the relationship between caloric intake and each of the
predictors changes.

```{r}
# Model 2 - Remove weight
healthmodel2 <- lm(Calories~Height+Protein+Carbohydrates+Calcium+
gender, data = healthdata)
summary(healthmodel2)
```
After removing weight, the p-values for all the predictor variables changed,
but the amounts of protein, carbohydrates, and calcium consumed by the subject
still remain the most significant predictors of their caloric intake.

```{r}
# Model 3 - Remove height
healthmodel3 <- lm(Calories~Protein+Carbohydrates+Calcium+
gender, data = healthdata)
summary(healthmodel3)
```
After removing height from our model, protein, carbohydrates, and calcium 
still remain the most significant predictors when working with a 95% confidence
level.

```{r}
# Model 4 - Remove gender
healthmodel4 <- lm(Calories~Protein+Carbohydrates+Calcium, data = healthdata)
summary(healthmodel4)
```
Now that all the non-significant predictors have been removed, we can determine
how well our final model fits the original data and if our model assumptions
are met.

```{r}
# Residual Plot - test to see if variances are equal
plot(healthmodel4$residuals ~ predict(healthmodel4), 
     main = "Residual Plot", xlab = "Fitted Values",
     ylab = "Residual Values")
```

```{r}
# Use shapiro.test() function to see if residuals are normally distributed
shapiro.test(healthmodel4$residuals)
```

Our final, simplified model helps us draw the following conclusions:

Starting at a value of approximately 27.09:

- For each unit increase in how much protein each subject consumes, their average caloric intake will increase by approximately 13.30.
- For each unit increase in how much carbohydrates the subject consumes, 
their average caloric intake will increase by approximately 4.63.
- For each unit increase in how much calcium each subject consumes, their
average caloric intake will decrease by approximately 0.23.
- Weight, height, and gender do not help with predicting the caloric intake
of a subject.

Since the p-value associated with the Shapiro-Wilk test is greater than .05, and
the graph labeled Residual Plot does not show any particular pattern (it looks like a random cloud centered at 0), this means that our model assumptions are satisfied. And because the value of R-squared (0.9295) is close to 1, this means that our final model fits the data very well and that a linear model was appropriate to use
for this analysis.

Therefore, we can also be certain that all of the conclusions which were drawn using
our final model are correct.
